---
title: NC groundwater well microbiomes

author: "Ariane L. Peralta, Randall Etheridge, Aied Garcia, Jake Hochard, Tom Vogel"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  fig_caption: yes
  pdf_document: null
header-includes:
- \usepackage{array}
- \usepackage{graphics}
- \usepackage[utf8]{inputenc}
---

Project Description: Fill out

# Initial Setup
```{r Initial Setup, include=FALSE}
rm(list=ls())
setwd("~/GitHub/GroundwaterMicrobiomes/analyses")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code Dependencies
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
require("vegan"); require("lme4"); require("reshape"); require("ggplot2")
require("ade4"); require("png"); require ("agricolae"); require("tidyr")
require("dplyr"); require("reshape2"); require("picante"); require("phyloseq")
```
#Import Files
## Environmental Data
```{r Import Files - Env, include=FALSE}
# Import Environmental Data
design <- read.csv("../data/GWQ_env.csv")
```
## Bacterial Data
```{r Import Files - Bacteria, include=FALSE}
# Import OTU data
# Import Raw Data
otu.in <- read.otu("../data/GWQ.bac.final.shared")

# Correct Sample IDs and Subset File
#missing <- setdiff(rownames(otu.in), rownames(design.2018))
#otu <- otu.in[-(which(rownames(otu.in) %in% missing)), ]
#dim(otu)
#otu <- otu[match(rownames(design.2018), rownames(otu)),]
#all.equal(rownames(otu), rownames(design.2018))
#dim(otu)

# OTU table - remove otus w/ < 2 occurrences across all sites
otus <- otu.in[, which(colSums(otu.in) >= 10)]
dim(otus)

# Make Presence Absence Matrix
dataPA <- (otus > 0) * 1

# Make Relative Abundance Matrices
dataREL <- otus
for(i in 1:dim(otus)[1]){
  dataREL[i,] <- otus[i,]/sum(otus[i,])
}

# Import Taxonomy File
otu.tax <- read.tax(taxonomy = "../data/GWQ.bac.final.taxonomy",
                   format = "rdp", tax.levels = 6, col.tax = 3)
```
## Subset E. coli positive only
```{r Ecoli positive, include=FALSE}
# combine design and micro dataframes
micro_env <- cbind(design,otus)

# subset TotCol_PA
ColPos <- subset(micro_env, TotCol_PA >=1)
dim(ColPos)
dim(design)
design <- (ColPos[,c(1:7)])
dim(design)
otus <- (ColPos[,-c(1:7)])
dim(otus)

dataREL <- otus
for(i in 1:dim(otus)[1]){
  dataREL[i,] <- otus[i,]/sum(otus[i,])
}

micro_env_sub <- cbind(design,dataREL)

# indicator for individual samples
library("labdsv")
design.type <- micro_env_sub$Timepoint
#design.type <- micro_env_sub[,c(3)] #dropped 
dataREL.t <- micro_env_sub[,-c(1:7)]
design.t <- micro_env_sub[,c(1:7)]
dataREL <- dataREL.t[, colSums(dataREL.t) > 0.05] # dominant subset
design.dataREL <- cbind(design.t,dataREL)

colnames <- colnames(dataREL)
tax.list <- as.data.frame(colnames)

ind.tax <- otu.tax[which(as.character(otu.tax$OTU) %in% tax.list$colnames), ]
ind.tax <- ind.tax[match(ind.tax$OTU, tax.list$colnames), ]

# Export OTU Table
write.table(ind.tax, "../data/DominantTaxa_ColPos.txt",
            sep="\t", row.names = F, quote = F)

# Export OTU Table
write.table(design.dataREL, "../data/dataREL_ColPos.txt",
            sep="\t", row.names = F, quote = F)
```
# Diversity Metrics - Hypothesis Testing
```{r Diversity Metrics - Bacteria, echo=TRUE}
# OTU table - removed low abundance samples
otu_low_remov <- otus[which(rowSums(otus) >= 1000), ]
dim(otu_low_remov)

rownames <- rownames(otus)
row.names(design) <- rownames
  
odd.sites <- otus[which(rowSums(otus) < 1000), ]
odd.sites <- rownames(odd.sites)

#aa <- (rowSums(otus))

# OTU table - odd sites in bacterial composition data and remove in design file
design_final <- design[setdiff(rownames(design), odd.sites), ]
all.equal(rownames(design_final), rownames(otu_low_remov))

min(rowSums(otu_low_remov))

otus <- otu_low_remov
```
# Diversity Metrics - Hypothesis Testing
```{r Diversity Metrics - Bacteria, echo=TRUE}
otus.r <- rrarefy(otus, 1000)

# Fisher's Alpha
fisher <- fisher.alpha(otus.r)

# Species Richness
#richness <- rowSums((PWESdata.r >= 1))
richness <- rowSums((otus >= 1))

# Shannon Diversity
shannon <- diversity(otus.r, "shannon")

# Simpson's Evenness
simp.even <- apply(otus.r, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(otus.r[,-c(1:1)]))

#combined richness, diversity, evenness
diversity <- cbind(design_final,richness,shannon,simp.even,J)
write.csv(diversity,"../data/diversity.bact.raw.csv")

diversity$TotCol_PA <- as.factor(diversity$TotCol_PA)

richness.lm <- lm(richness ~ TotCol_PA+Timepoint, diversity)
plot(richness.lm)
anova(richness.lm)

shannon.lm <- lm(shannon ~ TotCol_PA+Timepoint, diversity)
plot(shannon.lm)
anova(shannon.lm)

evenness.lm <- lm(simp.even ~ TotCol_PA+Timepoint, diversity)
plot(evenness.lm)
anova(evenness.lm)

# Graphing Shannon Diversity - Timepoint
p <- ggplot(diversity, aes(x=TotCol_PA, y=shannon, color=TotCol_PA))+ geom_boxplot() +
      geom_point(aes(color=factor(TotCol_PA)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Coliform Bacteria Status", values=c("gray70", "darkgreen"), labels = c("negative", "positive")) 
shannon<-p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("0","1"), labels=c("negative", "positive")) + theme(legend.position="none")
shannon
ggsave("../figures/shannon.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

# Graphing Richness
p <- ggplot(diversity, aes(x=TotCol_PA, y=richness, color=TotCol_PA))+ geom_boxplot() +
      geom_point(aes(color=factor(TotCol_PA)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Coliform Bacteria Status", values=c("gray70", "darkgreen"), labels = c("negative", "positive")) 
richness<-p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = "OTU Richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("0","1"), labels=c("negative", "positive")) + theme(legend.position="none")
richness
ggsave("../figures/richness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

# Graphing Simpson's Evenness
p <- ggplot(diversity, aes(x=TotCol_PA, y=simp.even, color=TotCol_PA))+ geom_boxplot() +
      geom_point(aes(color=factor(TotCol_PA)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Coliform Bacteria Status", values=c("gray70", "darkgreen"), labels = c("negative", "positive")) 
evenness<-p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Coliform Bacteria Status", y = "Simpson's Evenness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("0","1"), labels=c("negative", "positive")) + theme(legend.position="none")
evenness
ggsave("../figures/eveness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

# Graphing Shannon Diversity - Timepoint
p <- ggplot(diversity, aes(x=Timepoint, y=shannon, color=Timepoint)) + geom_boxplot() +
      geom_point(aes(color=factor(Timepoint)), size=2, position = position_jitterdodge())
shannon.tp <- p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) 
shannon.tp
#ggsave("../figures/shannon_TP.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

# Graphing Richness - Timepoint
p <- ggplot(diversity, aes(x=Timepoint, y=richness, color=Timepoint)) + geom_boxplot() +
      geom_point(aes(color=factor(Timepoint)), size=2, position = position_jitterdodge())
richness.tp<-p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = "OTU Richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1))
richness.tp
#ggsave("../figures/richness_TP.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

# Graphing Simpson's Evenness
p <- ggplot(diversity, aes(x=Timepoint, y=simp.even, color=Timepoint)) + geom_boxplot() +
      geom_point(aes(color=factor(Timepoint)), size=2, position = position_jitterdodge())
evenness.tp<-p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Coliform Bacteria Status", y = "Simpson's Evenness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("Florence_1_Dec2018","Florence_2_March2019","Florence_3_Aug2019","Florence_4_Sept2019"), labels=c("Dec.2018", "March.2019", "Aug.2019", "Sept.2019")) + theme(legend.position="none")
evenness.tp
#ggsave("../figures/eveness_TP.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
#combine alpha diversity graphs
```{r Loading Files, include=FALSE}
library("cowplot") #combining plots
library("ggpubr")
# remove legends and axes
n1 <-richness + theme(legend.position="top", legend.box = "horizontal")
n2 <- shannon + theme(legend.position="none")
n3 <- evenness + theme(legend.position="none")

# align all plots vertically
plots <- align_plots(n1,n2,n3, align = 'v', axis = 'l')

# combine moisture N process graphs
n4 <- plot_grid(plots[[1]],plots[[2]],plots[[3]], labels = c("A","B","C"), label_size = 16, ncol=1, align='v')
n4 

ggsave("../figures/alpha_diversity.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=9, height=9, dpi=300, limitsize=TRUE)
```
#combine alpha diversity graphs - timepoint
```{r Loading Files, include=FALSE}
# remove legends and axes
n5 <-richness.tp + theme(legend.position="top", legend.box = "horizontal")
n6 <- shannon.tp + theme(legend.position="none")
n7 <- evenness.tp + theme(legend.position="none")

# align all plots vertically
plots <- align_plots(n5,n6,n7, align = 'v', axis = 'l')

# combine moisture N process graphs
n8 <- plot_grid(plots[[1]],plots[[2]],plots[[3]], labels = c("A","B","C"), label_size = 16, ncol=1, align='v')
n8

ggsave("../figures/alpha_diversity.tp.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=9, height=9, dpi=300, limitsize=TRUE)
```

# Bacterial Ordinations - ColPos
```{r Ordination (PCoA) - Bacteria - bulk and teas, include=FALSE}
# Principal Coordinates Analysis
dataREL <- otus
for(i in 1:dim(otus)[1]){
  dataREL[i,] <- otus[i,]/sum(otus[i,])
}

dataREL2 <- dataREL[, colSums(dataREL) > 0.05] # dominant subset
dataREL.dist <- vegdist(dataREL2, method="bray")

pcoa <- cmdscale(dataREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #7.0
explainvar2b #5.6

# Principal Coordinates Analysis

pcoa.groups <- paste(design_final$TotCol_PA, sep = "_")
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
ColPos <- c("negative","positive")
pcoa.cent.dataframe.trts$ColPos <- as.factor(ColPos)

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=ColPos)) + theme_bw() + scale_color_manual(name="Coliform Bacteria Status", values=c("gray70", "darkgreen"), labels = c("negative", "positive")) 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
  geom_point(aes(colour=ColPos), size=5, stroke = 1.25, show.legend = TRUE) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (7.0%)") + ylab("PCoA 2 (5.6%)") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4))) 
ggsave("../figures/gwq_colpos_microbes.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=10, height=7, dpi=300, limitsize=TRUE)

#PERMANOVA
new.data <- cbind(design_final,dataREL2)
adonis = adonis(new.data[,-c(1:7)]~TotCol_PA, method = "bray", data = new.data, perm=1000, set.seed=42)
adonis

new.data <- cbind(design_final,dataREL2)
adonis = adonis(new.data[,-c(1:7)]~TotCol_PA+Timepoint, method = "bray", data = new.data, perm=1000, set.seed=42)
adonis

# Principal Coordinates Analysis all points plotted
pcoa.groups <- paste(design_final$TotCol_PA, sep = "_")
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

#Plot
plot1a <- ggplot(pcoa.points, aes(x=X1, y=X2, colour=group)) + theme_bw()+ scale_color_manual(name="Coliform Bacteria Status", values=c("gray70", "darkgreen"), labels = c("negative", "positive")) 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_point(aes(colour=group), size=4, stroke = 1.25, show.legend = TRUE) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (7.0%)") + ylab("PCoA 2 (5.6%)") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4))) 
ggsave("../figures/gwq_colpos_microbes_ALL.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=10, height=7, dpi=300, limitsize=TRUE)
```
# Bacterial community indicator species - ColPos
```{r Bacteria Indicator Species, include=FALSE}
library("labdsv")
#new.data <- cbind(design_final,dataREL2)
design.type <- new.data$TotCol_PA
dataREL.t <- new.data[,-c(1:7)]

dataREL <- dataREL.t[, colSums(dataREL.t) > 0.05]
bac.ind <- indval(dataREL, design.type)
levels(design.type)
summary(bac.ind)

inds <- which(bac.ind$pval <= 0.05)
bac.indicators <- as.data.frame(matrix(NA, nrow = length(inds), ncol = 4))
colnames(bac.indicators) <- c("OTU", "Cluster", "IndVal", "Prob")

bac.indicators$OTU <- names(inds)
bac.indicators$Cluster <- bac.ind$maxcls[inds]
bac.indicators$IndVal <- bac.ind$indcls[inds]
bac.indicators$Prob <- bac.ind$pval[inds]

ind.tax <- otu.tax[which(as.character(otu.tax$OTU) %in% bac.indicators$OTU), ]
ind.tax <- ind.tax[match(ind.tax$OTU, bac.indicators$OTU), ]

indicator.bac <- cbind(bac.indicators, ind.tax[, -c(1)])

indicator.bac <- indicator.bac[order(as.numeric(indicator.bac$Cluster)), ]

table(indicator.bac$Cluster)
table(indicator.bac$Phylum)
table(indicator.bac$Cluster)
levels(design.type)

# Export Bacteria Indicator Table
write.table(indicator.bac, "../data/BacterialIndicators_ColPos.txt",
            sep="\t", row.names = F, quote = F)
```
# Bacterial Ordinations - Timepoints
```{r Ordination (PCoA) - Bacteria - bulk and teas, include=FALSE}
# Principal Coordinates Analysis
pcoa.groups <- paste(new.data$Timepoint, sep = "_")
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
Timepoint <- c("Florence_1_Dec2018","Florence_2_March2019","Florence_3_Aug2019","Florence_4_Sept2019")
pcoa.cent.dataframe.trts$Timepoint <- as.factor(Timepoint)

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=Timepoint)) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
    geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
    geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
    geom_point(aes(colour=Timepoint), size=5, stroke = 1.25, show.legend = TRUE) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (7.0%)") + ylab("PCoA 2 (5.6%)") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4))) 
ggsave("../figures/gwq_microbes_Timepoint.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=10, height=7, dpi=300, limitsize=TRUE)
```
# Bacterial community indicator species
```{r Bacteria Indicator Species, include=FALSE}
library("labdsv")
new.data <- cbind(design_final,dataREL2)
design.type <- new.data$Timepoint
design.type <-droplevels(new.data$Timepoint)
dataREL.t <- new.data[,-c(1:7)]

dataREL2 <- dataREL.t[, colSums(dataREL.t) > 0.05]
#Timepoint <- c("Florence_1_Dec2018","Florence_2_March2019","Florence_3_Aug2019","Florence_4_Sept2019")
bac.ind <- indval(dataREL2, design.type)

summary(bac.ind)

inds <- which(bac.ind$pval <= 0.05)
bac.indicators <- as.data.frame(matrix(NA, nrow = length(inds), ncol = 4))
colnames(bac.indicators) <- c("OTU", "Cluster", "IndVal", "Prob")

bac.indicators$OTU <- names(inds)
bac.indicators$Cluster <- bac.ind$maxcls[inds]
bac.indicators$IndVal <- bac.ind$indcls[inds]
bac.indicators$Prob <- bac.ind$pval[inds]

ind.tax <- otu.tax[which(as.character(otu.tax$OTU) %in% bac.indicators$OTU), ]
ind.tax <- ind.tax[match(ind.tax$OTU, bac.indicators$OTU), ]

indicator.bac <- cbind(bac.indicators, ind.tax[, -c(1)])

indicator.bac <- indicator.bac[order(as.numeric(indicator.bac$Cluster)), ]

table(indicator.bac$Cluster)
table(indicator.bac$Phylum)
table(indicator.bac$Cluster)
levels(design.type)

# Export Bacteria Indicator Table
write.table(indicator.bac, "../data/BacterialIndicators_Timepoint.txt",
            sep="\t", row.names = F, quote = F)
```
# Simple Hypothesis Testing - Microbes
```{r redundancy analysis - Bacteria, echo=TRUE}
# RDA (full data - works)
new.data <-cbind(design,dataREL)
rda.2 <- rda(decostand(otus, "hell") ~  nitrate+Timepoint+TotCol_PA, data=design, na=na.omit)
RsquareAdj(rda.2)

# TEST SUBSET DATA (works)
otus.sub <- otus[(1:10),(1:10)]
design.sub <- design[(1:10),]
dataREL.sub <- otus.sub
for(i in 1:dim(otus.sub)[1]){
  dataREL.sub[i,] <- otus.sub[i,]/sum(otus.sub[i,])
} 

rda(dataREL.sub ~ nitrate, data=design.sub, na=na.omit)

new.data <-cbind(design.sub,dataREL.sub)
adonis = adonis(new.data[,-c(1:7)]~nitrate, method = "bray", data = new.data, perm=1000)
adonis

rda(otus.sub ~ nitrate, data=design.sub)

rda.1 <- rda(decostand(otus.sub, "hell") ~  nitrate, data=design.sub)
RsquareAdj(rda.1)
```